---
title: "Final_effect_size_human_chimp_conservation_status"
author: "Lauren Blake"
date: "June 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
getwd()
# import sample labels
samples <- read.delim("../../../Reg_Evo_Primates/data/Sample_info_RNAseq_RIN.txt")
RIN <- samples$RIN
# expression
exprs <- read.table("../../../Reg_Evo_Primates/data/human_chimp_orth_exp_methyl_7725_hum.txt", sep="")
# methylation data
methyl <- read.csv("../../../Reg_Evo_Primates/data/chimp_human_orth_7725_avg_methyl_per_ts_gene.txt", sep="")

# Normalized gene expression data
cpm.voom.cyclic <- readRDS("../../../Reg_Evo_Primates/data/human_chimp_orth_cpm_voom_cyclic.rds")


# Load libraries
library("edgeR")
library("limma")
library("plyr")
library("ashr")
library("cowplot")
library("vashr")
library("devtools")
#devtools::install_github("jhsiao999/mediation/pkg")
library("medinome")
```

# Set FDR level/svalue

```{r}
FDR_level <- 0.05
#FDR_add <- 0.10
```

# Chimp heart-kidney

```{r}
# Check parameters

human_chimp_heart <- c(1, 2, 5, 6, 9, 10, 13, 14)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]


methyl <- exprs_methyl[,human_chimp_heart]
rownames(hc_exprs_heart) <- rownames(exprs)
rownames(methyl) <- rownames(exprs)


Y <- hc_exprs_heart
two_species <- samples$Tissue[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]



check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash_normal <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")
summary(fit_ash_normal$result$svalue < 0.05)


```



## Human heart versus kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human heart versus kidney- conserved DE genes

```{r}
# Species no DE
species_no_DE <- c(17, 20, 21, 24, 25, 28, 29)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] <  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

```{r}
# Species no DE
species_no_DE <- c(17, 20, 21, 24, 25, 28, 29)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] >  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human heart versus kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Chimp heart-liver

```{r}
# Check parameters

human_chimp_heart <- c(1, 3, 5, 7, 9, 11, 13, 15)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]


methyl <- exprs_methyl[,human_chimp_heart]
rownames(hc_exprs_heart) <- rownames(exprs)
rownames(methyl) <- rownames(exprs)


Y <- hc_exprs_heart
two_species <- samples$Tissue[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]



check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash_normal <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")
summary(fit_ash_normal$result$svalue < 0.05)


```



## Human heart versus lung- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human heart versus kidney- conserved DE genes

```{r}
# Species no DE
species_no_DE <- c(18, 20, 22, 24, 26, 28, 30)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] <  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

```{r}
# Species no DE
species_no_DE <- c(18, 20, 22, 24, 26, 28, 30)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] >  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human heart versus kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


# Chimp heart-lung

```{r}
# Check parameters

human_chimp_heart <- c(1, 4, 5, 8, 9, 12, 13, 16)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]


methyl <- exprs_methyl[,human_chimp_heart]
rownames(hc_exprs_heart) <- rownames(exprs)
rownames(methyl) <- rownames(exprs)


Y <- hc_exprs_heart
two_species <- samples$Tissue[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]



check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash_normal <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")
summary(fit_ash_normal$result$svalue < 0.05)


```



## Human heart versus kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human heart versus lung- conserved DE genes

```{r}
# Species no DE
species_no_DE <- c(19, 20, 23, 24, 27, 28, 31)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] <  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

```{r}
# Species no DE
species_no_DE <-  c(19, 20, 23, 24, 27, 28, 31)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] >  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human heart versus kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


# Chimp kidney-liver

```{r}
# Check parameters

human_chimp_heart <- c(2, 3, 6, 7, 10, 11, 14, 15)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]


methyl <- exprs_methyl[,human_chimp_heart]
rownames(hc_exprs_heart) <- rownames(exprs)
rownames(methyl) <- rownames(exprs)


Y <- hc_exprs_heart
two_species <- samples$Tissue[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]



check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash_normal <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")
summary(fit_ash_normal$result$svalue < 0.05)


```



## Human heart versus kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human heart versus lung- conserved DE genes

```{r}
# Species no DE
species_no_DE <- c(17, 18, 21, 22, 25, 26, 29, 30)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] <  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

```{r}
# Species no DE
species_no_DE <-  c(17, 18, 21, 22, 25, 26, 29, 30)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] >  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human heart versus kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


# Chimp kidney-lung

```{r}
# Check parameters

human_chimp_heart <- c(2, 4, 6, 8, 10, 12, 14, 16)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]


methyl <- exprs_methyl[,human_chimp_heart]
rownames(hc_exprs_heart) <- rownames(exprs)
rownames(methyl) <- rownames(exprs)


Y <- hc_exprs_heart
two_species <- samples$Tissue[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]



check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash_normal <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")
summary(fit_ash_normal$result$svalue < 0.05)


```



## Human heart versus kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human heart versus lung- conserved DE genes

```{r}
# Species no DE
species_no_DE <- c(17, 19, 21, 23, 25, 27, 29, 31)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] <  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

```{r}
# Species no DE
species_no_DE <-  c(17, 19, 21, 23, 25, 27, 29, 31)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] >  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human heart versus kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


# Chimp liver-lung

```{r}
# Check parameters

human_chimp_heart <- c(3, 4, 7, 8, 11, 12, 15, 16)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]


methyl <- exprs_methyl[,human_chimp_heart]
rownames(hc_exprs_heart) <- rownames(exprs)
rownames(methyl) <- rownames(exprs)


Y <- hc_exprs_heart
two_species <- samples$Tissue[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]



check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash_normal <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")
summary(fit_ash_normal$result$svalue < 0.05)


```



## Human heart versus kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human heart versus lung- conserved DE genes

```{r}
# Species no DE
species_no_DE <- c(18, 19, 22, 23, 26, 27, 30, 31)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] <  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

```{r}
# Species no DE
species_no_DE <-  c(18, 19, 22, 23, 26, 27, 30, 31)

species_no_DE_exprs <- hc_exprs[,species_no_DE]

species_no_DE_methyl <- exprs_methyl[,species_no_DE]

two_species <- samples$Tissue[species_no_DE]
no_DE_tissue <- droplevels.factor(two_species)

no_DE_RIN <- RIN[species_no_DE]

design <- model.matrix(~ as.factor(no_DE_tissue) + no_DE_RIN)
  fit_species_no_DE <- lmFit(cpm.voom.cyclic[,species_no_DE], design)
  fit_species_no_DE <- eBayes(fit_species_no_DE)

  HvC_Heart_fit_species_no_DE = topTable(fit_species_no_DE, coef=2, adjust="BH", number=Inf,
sort.by="none")

# Species DE 
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  
  HvC_Heart_fit_species_no_DE = cbind(HvC_Heart_fit_species_no_DE, HvC_Heart_fit_all)
  
 
      HvC_Heart_fit_all_5perc <- HvC_Heart_fit_species_no_DE[which(HvC_Heart_fit_species_no_DE[,6] >  FDR_level & HvC_Heart_fit_species_no_DE[,13] < FDR_level), ]

  
   human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

summary(fit_ash$result$svalue < FDR_level)
calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human heart versus kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```
