---
title: "Effect_size_updated_2lms"
author: "Lauren Blake"
date: "June 18, 2018"
output: html_document
---

This is a script to evaluate the effect sizes of the 2 linear models. 

## Load in data 


```{r chunk-options, include=FALSE}
source("chunk-options.R")
```


```{r}
getwd()
# import sample labels
samples <- read.delim("../../../Reg_Evo_Primates/data/Sample_info_RNAseq_RIN.txt")
RIN <- samples$RIN
# expression
exprs <- read.table("../../../Reg_Evo_Primates/data/250_exp_avg_methyl_hcr_4155_genes.txt", sep="")


# Normalized gene expression data
cpm.voom.cyclic <- readRDS("../../../Reg_Evo_Primates/data/human_chimp_orth_cpm_voom_cyclic.rds")


# Load libraries
library("edgeR")
library("limma")
library("plyr")
library("ashr")
library("cowplot")
library("vashr")
library("devtools")
#devtools::install_github("jhsiao999/mediation/pkg")
library("medinome")
```

# Set FDR level/svalue

```{r}
FDR_level <- 0.05
```

# Human-chimp heart

```{r}
# Check parameters

human_chimp_heart <- c(1, 5, 9, 13, 20, 24, 28)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]


Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]



check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=5, singlecomp = T, unimodal = "auto") 
fit_ash_normal <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")
summary(fit_ash_normal$result$svalue < 0.05)


```



## Human versus chimpanzee heart- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=5, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human versus chimpanzee heart- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=5, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp kidney

```{r}
# Check parameters

human_chimp_heart <- c(2, 6, 10, 14, 17, 21, 25, 29)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]



Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp liver

```{r}
# Check parameters

human_chimp_heart <- c(3, 7, 11, 15, 18, 22, 26, 30)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]



Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee liver- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee liver- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp lung

```{r}
# Check parameters

human_chimp_heart <- c(4, 8, 12, 16, 19, 23, 27, 31)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only

hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]


Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee lung- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee lung- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Rerun at FDR/FSR = 0.1

```{r}
FDR_level <- 0.1
```

# Human-chimp heart

```{r}
# Check parameters

human_chimp_heart <- c(1, 5, 9, 13, 20, 24, 28)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only

hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]



Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]


```



## Human versus chimpanzee heart- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=5, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human versus chimpanzee heart- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=5, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp kidney

```{r}
# Check parameters

human_chimp_heart <- c(2, 6, 10, 14, 17, 21, 25, 29)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]


Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp liver

```{r}
# Check parameters

human_chimp_heart <- c(3, 7, 11, 15, 18, 22, 26, 30)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]


Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee liver- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee liver- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp lung

```{r}
# Check parameters

human_chimp_heart <- c(4, 8, 12, 16, 19, 23, 27, 31)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only

hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]



Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee lung- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee lung- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Rerun at FDR/FSR = 0.01

```{r}
FDR_level <- 0.01
```

# Human-chimp heart

```{r}
# Check parameters

human_chimp_heart <- c(1, 5, 9, 13, 20, 24, 28)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only

hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]


Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]


```



## Human versus chimpanzee heart- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)
fit <- vash(check_values$d_se,df=5, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```


## Human versus chimpanzee heart- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:7]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=5, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp kidney

```{r}
# Check parameters

human_chimp_heart <- c(2, 6, 10, 14, 17, 21, 25, 29)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]



Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee kidney- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee kidney- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp liver

```{r}
# Check parameters

human_chimp_heart <- c(3, 7, 11, 15, 18, 22, 26, 30)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only


hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]



Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee liver- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee liver- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

# Human-chimp lung

```{r}
# Check parameters

human_chimp_heart <- c(4, 8, 12, 16, 19, 23, 27, 31)

# Methylation
hc_exprs <- exprs[,2:48]
exprs_methyl <- exprs[,49:79]

# HC heart only

hc_exprs_heart <- hc_exprs[,human_chimp_heart]
rownames(hc_exprs_heart) <- exprs[,1]

methyl <- exprs_methyl[,human_chimp_heart]
rownames(methyl) <- exprs[,1]



Y <- hc_exprs_heart
two_species <- samples$Species[human_chimp_heart]
X <- droplevels.factor(two_species)
M <- methyl
RIN_subset <- RIN[human_chimp_heart]

```

## Human versus chimpanzee liver- DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```

## Human versus chimpanzee liver- non DE genes

```{r}
# Run the linear model in limma
  design_1 <- model.matrix(~ X + RIN_subset)
  fit_all <- lmFit(cpm.voom.cyclic[,human_chimp_heart], design_1)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val > FDR_level), ]
  
 human_chimp_heart1 <-  rownames(methyl) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(methyl, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
M <- counts_human_chimp_heart_subset

human_chimp_heart1 <-  rownames(hc_exprs_heart) %in% rownames( HvC_Heart_fit_all_5perc)
human_chimp_heart1 <- as.data.frame(human_chimp_heart1)
counts_genes_in <- cbind(hc_exprs_heart, human_chimp_heart1)
counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart1 == "TRUE")
counts_human_chimp_heart_subset <- counts_genes_in_cutoff[,1:8]
Y <- counts_human_chimp_heart_subset

check_values <- mediate.test.regressing(Y, X, M, RIN_subset)

fit <- vash(check_values$d_se,df=6, singlecomp = T, unimodal = "auto") 
fit_ash <- ash(as.vector(check_values$d), fit$sd.post, mode = 0, mixcompdist = "normal")

calc <- as.array(summary(fit_ash$result$svalue < FDR_level))
calc2 <- as.numeric(calc[[2]])

(nrow(Y)-calc2)/nrow(Y)
```